### 2018-04-11 paper/overview.tex

% solving the running example
\subsection{$\mcchc$ as a Clause-Dependence Disjoint System}
\label{sec:solve-ex}
% state it is CDD
The recursion-free CHC system $\mcchc$ is a
\emph{Clause-Dependence-Disjoint} (CDD) system.
% state the defn of CDD:
A CHC system can be classified as CDD when the clauses obey
rules based on dependencies between the predicates.
%
One predicate is directly dependent on another if the second appears
in the body of a clause where the first is the head.
%
Based on this, we say that a clause is dependence disjoint if
\textbf{(1)} no two predicates in the body share any transitive
dependencies on other predicates and \textbf{(2)} no two predicates in
the body are the same.
% give a example
As an example, clause $(7)$ is dependence disjoint. Clause $(7)$ uses
both $\cc{L}_9$ and $\cc{dbl}$ in its body.
%
The transitive dependency of $\cc{L}_9$ is the set
$\{\cc{L}_4,\cc{L}_6,\cc{L}_8\}$ while the transitive dependency of
\cc{dbl} is the empty set.
%
Therefore, their transitive dependencies are disjoint:
$\{\cc{L}_4,\cc{L}_6,\cc{L}_8\} \cap \varnothing = \varnothing$.
%
All other clauses in $\mcchc$ have at most one uninterpreted predicate
in the body, so they are trivially disjoint dependent.
%
When each clause in a CHC system is dependence disjoint, it is a CDD
system. Therefore $\mcchc$ is a CDD system.
%
%
The formal definition of CDD and its key properties are given in
\autoref{sec:CDD-defn}.
The formal definition of transitive dependency is given in\QZ{reference}.
%

% restate defn of CDD systems:
\sys solves CDD systems directly by issuing a binary interpolation
query for each uninterpreted predicate in topological order.
%
Each uninterpreted predicate $P$ has a valid interpretation which can
be computed by interpolating
\textbf{(1)} the \emph{pre}-formula, constructed from clauses where
$P$ is the head and
\textbf{(2)} the \emph{post}-formula, constructed from
all clauses where the head transitively depends on $P$.
%

We will walk through the construction of an interpolation query for
$\cc{L}_9$ by hand.  By the time \sys reaches $\cc{L}_9$ it will
already have solutions for $\cc{L}_4$, $\cc{L}_6$, $\cc{L}_8$.
Possible interpretations of these predicates are shown in
\autoref{fig:ex-graph}.
%
A solution for $\cc{L}_9$ can be constructed by interpolating
the appropriate pre and post formulas.
%
The pre-formula is constructed from the bodies of clauses where
$\cc{L}_9$ is the head. Each relational
predicate, $P$,
is replaced by a corresponding boolean indicator variable, $\cc{b}_P$, and each
boolean indicator variable implies the solution for its predicate.
%
In particular, the pre-formula for $\cc{L}_9$ is constructed from
clauses (5) and (6):
\begin{gather}
  ((\cc{b}_{\cc{L}_6} \land \cc{abs'} = \cc{n})
  \lor
  (\cc{b}_{\cc{L}_8} \land \cc{abs'} = -\cc{n}))
  \land
  (\cc{b}_{\cc{L}_6} \Rightarrow \cc{n} \ge 0)
  \land
  (\cc{b}_{\cc{L}_8} \Rightarrow \cc{n} < 0)
\end{gather}
%
The post-formula is constructed from clauses that transitively depend
on $\cc{L}_9$. Again, we replace relational predicates by
corresponding boolean indicators. However, we omit the boolean indicator for
$\cc{L}_9$. The post-formula is composed from clauses (1), (7), and
(8):
\begin{gather}
  (\cc{b}_{\cc{main}} \land \cc{res'} < 0)
  \land
  (\cc{b}_{\cc{main}} \Rightarrow
    (\cc{b}_{\cc{dbl}}
    \land \cc{abs'}= \cc{x}
    \land \cc{res}=\cc{d} ))
  \land (\cc{b}_{\cc{dbl}} \Rightarrow \cc{d} = 2*\cc{x})
\end{gather}
%
Interpolating the pre and post formulas yields a valid intepretation
of $\cc{L}_9$: $\cc{abs'} \geq 0$.
The procedure for solving a CDD system is described in formal detail
in \autoref{sec:solve-cdd}.

\subsection{$\mcchc$ is not in other recursion-free classes}
\label{sec:not-in}
%
In this section, we show that $\mcchc$ cannot be placed in other
known classes of recursion-free CHC systems. Specifically, we will
discuss body-disjoint systems and linear systems.
%

% tree solver
Tree structure
systems~\cite{mcmillan14,bjorner13,heizmann10,rummer13a,rummer13b} are
class of recursion-free CHC system where each uninterpreted predicate
appears in the body of at most one clause (and appears in such a
clause exactly once).
%
Such systems cannot model a program with multiple control paths that
share a common subpath, typically modeled as a CHC system with a
uninterpreted predicate that occurs in the body of multiple clauses.
%
$\mcchc$ is not a body-disjoint system because the uninterpreted
predicates $\cc{L}_4$ shows in the body of both clause (3) and clause
(4).
%
In order to find an interpretation for $\mcchc$, a solver which uses
tree-structure systems would have to create a fresh copy of
$\cc{L}_4$, as well copies for all relational predicates that
$\cc{L}_4$ depends on.

@@

The class CDD of CDD systems is a superclass of the classes of
body-disjoint and linear systems.
%
% The fact that $\mcchc$ is CDD corresponds to the fact that there no
% statements executed more than once within one execution path.
% \DAH{elaborate}
%
E.g., a recursion-free system that models \cc{dblAbs} 
call \cc{dbl} twice on \cc{abs} in sequence would not
be CDD.
%
\sys, given an arbitrary recursion-free CHC system $S$,
solves it by generating an equivalent CDD system $S'$ and
solving $S'$ directly.
%
In general, $S'$ may have size exponential in the size of
$S$.
%
In practice, a suitable equivalent system can be generated using
heuristics analogous to those used to generate compact verification
conditions of hierarchical programs~\cite{flanagan01,lal-qadeer15}.
%
A general procedure for constructing a CDD expansion of a given CHC
system is given in \autoref{app:cons-cdd}.

>>>>

% solving the running example
\subsection{$\mcchc$ as a Clause-Dependence Disjoint System}
\label{sec:solve-ex}
% state it is CDD
The recursion-free CHC system $\mcchc$ is a
\emph{Clause-Dependence-Disjoint} (CDD) system.
% state the defn of CDD:
A CHC system can be classified as CDD when all clauses are satisified
the following rules:
%
\textbf{(1)} no two predicates in the body of the same clause share any transitive
dependencies on other predicates(i.e. dependence disjoint) 
and \textbf{(2)} no clause has more than one same predicate in its body.
% give a example
As an example, clause $(7)$ is dependence disjoint. 
%
Two predicates $\cc{L}_9$ and $\cc{dbl}$ are in its body.
%
The transitive dependency of $\cc{L}_9$ is the set
$\{\cc{L}_4,\cc{L}_6,\cc{L}_8\}$ while the transitive dependency of
\cc{dbl} is the empty set.
%
Therefore, their transitive dependencies are disjoint:
$\{\cc{L}_4,\cc{L}_6,\cc{L}_8\} \cap \varnothing = \varnothing$.
%
All other clauses in $\mcchc$ have at most one uninterpreted predicate
in the body, so they are trivially disjoint dependent.
%
And there is no clauses has more than one same predicate in its body.
%
Therefore $\mcchc$ is a CDD system.
%
%
The formal definition of CDD and its key properties are given in
\autoref{sec:CDD-defn}.
The formal definition of transitive dependency is given in
\QZ{reference}.
%

% restate defn of CDD systems:
\sys solves CDD systems directly by issuing a binary interpolation
query for each uninterpreted predicate in topological order.
%
Each interpretation of uninterpreted predicate $P$ can
be computed by interpolating
\textbf{(1)} the \emph{pre}-formula, constructed from clauses where
$P$ is the head and
\textbf{(2)} the \emph{post}-formula, constructed from
all clauses where the head transitively depends on $P$.
%

For example, when \sys try to solve $\cc{L}_9$, it has partial solution that 
map $\cc{L}_4$, $\cc{L}_6$, $\cc{L}_8$ to their interpretation.
%
Possible interpretations of these predicates are shown in
\autoref{fig:ex-graph}.
%

%
A interpretation of $\cc{L}_9$ can be constructed by interpolating
the appropriate pre and post formulas.
%
The pre-formula is constructed from the bodies of clauses where
$\cc{L}_9$ is the head. 
%
Each relational predicate, $P$,
is replaced by a corresponding boolean indicator variable, $\cc{b}_P$.
%
Each
boolean indicator variable implies the solution for its predicate which is encoded
as the disjunction of the negation of the boolean indicator variable and the solution.
%
In particular, the pre-formula for $\cc{L}_9$ is constructed from
clauses (5) and (6):
\begin{gather}
  ((\cc{b}_{\cc{L}_6} \land \cc{abs'} = \cc{n})
  \lor
  (\cc{b}_{\cc{L}_8} \land \cc{abs'} = -\cc{n}))
  \land
  (\neg \cc{b}_{\cc{L}_6} \lor \cc{n} \ge 0)
  \land
  (\neg \cc{b}_{\cc{L}_8} \lor \cc{n} < 0)
\end{gather}
%
The post-formula is constructed from clauses that transitively depend
on $\cc{L}_9$. Again, we replace relational predicates by
corresponding boolean indicators. However, we omit the boolean indicator for
$\cc{L}_9$. The post-formula is composed from clauses (1), (7), and
(8):
\begin{gather}
  (\cc{b}_{\cc{main}} \land \cc{res'} < 0)
  \land
  (\neg \cc{b}_{\cc{main}} \lor
    (\cc{b}_{\cc{dbl}}
    \land \cc{abs'}= \cc{x}
    \land \cc{res}=\cc{d} ))
  \land (\neg \cc{b}_{\cc{dbl}} \lor \cc{d} = 2*\cc{x})
\end{gather}
%
Interpolating the pre and post formulas yields a intepretation
of $\cc{L}_9$: $\cc{abs'} \geq 0$.
%
The procedure for solving a CDD system is described in formal detail
in \autoref{sec:solve-cdd}.

\subsection{$\mcchc$ is not in other recursion-free classes}
\label{sec:not-in}
%
In this section, we show that $\mcchc$ is not in other
known classes of recursion-free CHC systems. 
%
Specifically, we will
discuss body-disjoint systems and linear systems.
%

% tree solver
Body-disjoint systems(i.e derivation tree)
systems~\cite{mcmillan14,bjorner13,heizmann10,rummer13a,rummer13b} are
class of recursion-free CHC system where each uninterpreted predicate
appears in the body of at most one clause and appears in such a
clause exactly once.
%
Such systems cannot model a program with multiple control paths that
share a common pre-subpath, typically modeled as a CHC system with a
uninterpreted predicate that occurs in the body of multiple clauses.
%
$\mcchc$ is not a body-disjoint system because the uninterpreted
predicates $\cc{L}_4$ shows in the body of both clause (3) and clause
(4).
%
In order to find an interpretation for $\mcchc$, a solver which uses
tree-structure systems would have to create a fresh copy of
$\cc{L}_4$, as well copies for all relational predicates that
$\cc{L}_4$ depends on.

@@

The class CDD of CDD systems is a superclass of union of the classes of
body-disjoint and linear systems.
%
% The fact that $\mcchc$ is CDD corresponds to the fact that there no
% statements executed more than once within one execution path.
% \DAH{elaborate}
%
E.g., a recursion-free system that models \cc{dblAbs} 
call \cc{dbl} twice on \cc{abs} in sequence would not
be CDD.
%
\sys, given an arbitrary recursion-free CHC system $S$,
solves it by generating an equivalent CDD system $S'$ and
solving $S'$ directly.
%
In general, $S'$ may have size exponential in the size of
$S$.
%
In practice, a suitable equivalent system can be generated using
heuristics analogous to those used to generate compact verification
conditions of hierarchical programs~\cite{flanagan01,lal-qadeer15}.
%
A general procedure for constructing a CDD expansion of a given CHC
system is given in \autoref{app:cons-cdd}.
